Time and space specifications for stencils

### 2016-06-19 paper.tex
typo
redoing the equational theory

\newcommand{\stenFwd}[2]{\texttt{forward} \, (\texttt{depth=}#1,
  \texttt{dim=}#2)}
\newcommand{\stenBwd}[2]{\texttt{backward} \, (\texttt{depth=}#1,
  \texttt{dim=}#2)}
\newcommand{\stenCen}[2]{\texttt{centered} \, (\texttt{depth=}#1,
  \texttt{dim=}#2)}

@@

\begin{figure}[t]
\begin{align*}
\def\arraystretch{1.2}
\setlength{\arraycolsep}{0.2em}
\newcommand{\dimTy}{\mathbb{D}}
\begin{array}{rl}
\nonterm{specification} ::= & \nonterm{regionDec} \mid \nonterm{specDec} \\
\nonterm{specDec} ::= & \term{stencil} \; \nonterm{spec}
  \\
\nonterm{regionDec} ::= &  \texttt{region} \; \nonterm{rvar} \; \texttt{=} \;
                         \nonterm{region}\\
\nonterm{spec} ::= & \nonterm{spatial} \mid \nonterm{temporal}   \\[1em]
\nonterm{spatial} ::= & [\nonterm{mod}] \;
                       [\nonterm{approxMod},] \; [\nonterm{region}] \\
\nonterm{mod} ::= & [\nonterm{modR},] \; [\nonterm{modRefl},] \; [\nonterm{modIrefl},] \\
\nonterm{modR} ::= &
 \term{readOnce} \\
\nonterm{modRefl} ::= & \term{reflexive}, \; \term{dims=}\dimTy \{\term{,} \; \dimTy\} \\
\nonterm{modIrrefl} ::= & \term{irreflexive}, \; \term{dims=}\dimTy \{\term{,}
          \dimTy\}\\
\nonterm{approxMod} ::= & \term{atMost} \; \mid \; \term{atLeast} \\[0.1em]
\nonterm{region} ::= & \term{forward}, \term{depth=}\mathbb{N}, \term{dim=}\dimTy  \\
\mid \; & \term{backward}, \term{depth=}\mathbb{N}, \term{dim=}\dimTy \\
\mid \; & \term{centered}, \term{depth=}\mathbb{N}, \term{dim=}\dimTy \\
\mid \; & \nonterm{region} \; \term{+} \; \nonterm{region} \\
\mid \; & \nonterm{region} \; \term{*} \; \nonterm{region} \\
\mid \; & \nonterm{rvar}  \\[0.5em]
\nonterm{temporal} ::= \; & \term{dependency} \; (v \; \{ , v \})
  \\[0.5em]
\dimTy ::= \; & \mathbb{N}_{>0} \\
\nonterm{rvar} ::= \; & [\text{\term{a}-\term{z}$\,$\term{A}-\term{Z}$\,$\term{0}-\term{9}}]+
\end{array}
\end{align*}
\caption{Specification syntax (EBNF grammar)}
\label{fig:syntax}
\end{figure}

@@

\begin{figure}
\begin{align*}
\hspace{-1em}
\setlength{\arraycolsep}{0.05em}
\begin{array}{c}
\framebox{$\overline{\texttt{v} : S} \equiv \overline{\texttt{u} : T}$} \\[1em]
\begin{array}{rl}
(\textsc{coalesceF}) \;\; & (\texttt{a} : \texttt{fwd} \;
                       \texttt{dims}=ds \; \texttt{depth}=n; \\
&  \;\, \texttt{a} :  \texttt{fwd} \; \texttt{dims}=ds \; \texttt{depth}=n+1)\\
\equiv \; & \texttt{a} :  \texttt{fwd} \; \texttt{depth}=n+1,
         \texttt{dims}=ds \\[0.75em]
(\textsc{coalesceB}) \;\; & (\texttt{a} : \texttt{bwd} \;
                       \texttt{dims}=ds \; \texttt{depth}=n; \\
&  \;\, \texttt{a} :  \texttt{bwd} \; \texttt{dims}=ds \; \texttt{depth}=n+1) \\
\equiv \; & \texttt{a} :  \texttt{bwd} \; \texttt{depth}=n+1,
         \texttt{dims}=ds \\[0.75em]
(\textsc{idem}) \;\; & (\texttt{a} : S; \texttt{a} : S) \equiv (\texttt{a}
                  : S)
\end{array} \\ \\
\framebox{$\texttt{v} : S \equiv \texttt{v'} : S'$} \\[1em]
\begin{array}{ll}
(\texttt{a} : \; \texttt{bwd} \; \texttt{depth=1\{b\}} \;
  \texttt{dim=t})
\, & \equiv \,
(\texttt{b} : \; \texttt{bwd} \; \texttt{depth=1\{a\}} \;
  \texttt{dim=t})
\\
(\texttt{a} : \; \texttt{fwd} \; \texttt{depth=1\{b\}}\;
  \texttt{dim=t})
\, & \equiv \,
(\texttt{b} : \; \texttt{fwd} \; \texttt{depth=1\{a\}} \;
  \texttt{dim=t})
\end{array}
\\ \\
\framebox{$S \equiv S'$} \\[0.75em]
\begin{array}{ll}
(a : \texttt{read-once} \; S) \; & \equiv \; (a : \texttt{read-once} \;\,
  \texttt{read-once} \; \, S)
\end{array}
\end{array}
\end{align*}
\caption{Equations on specifications}
\label{fig:equations}
\end{figure}

@@

%% VECTOR NOTATIONS
\newcommand{\vect}[1]{\textbf{#1}}
\newcommand{\vtwo}[2]{\setlength{\arraycolsep}{0em}
\left[\begin{array}{l}$#1$\\$#2$\end{array}\right]}
\newcommand{\stwo}[4]
%{\vtwo{#1}{#2}\!\vtwo{#3}{#4}}
{\setlength{\arraycolsep}{0.1em}
\left[\begin{array}{rr}$#1$ & $#3$\\$#2$ & $#4$\end{array}\right]}

@@

\item Sort each permutation set into an ordered list, based on the ordering:
\[
\vect{x} \leq \vect{y} = \exists i \, . \; \vect{x}_i \leq \vect{y}_i 
\; \wedge \; \vect{x}_{i:n} = \vect{y}_{i:n}
\]
%
This yields:
%%
\begin{align*}
\{&(\pi^2_1, [
\stwo{0}{-1}{0}{-1}
\stwo{-1}{0}{-1}{0} 
\stwo{0}{0}{0}{0}
\stwo{1}{0}{1}{0} 
\stwo{0}{1}{0}{1}] )
\\
&(\pi^2_2, [
\stwo{0}{-1}{0}{-1}
\stwo{-1}{0}{-1}{0}
\stwo{0}{0}{0}{0}
\stwo{1}{0}{1}{0}
\stwo{0}{1}{0}{1}])\}
\end{align*}
%%
\item Fold each list pairwise by the following partial $\bullet$ which 
coalescees and contiguous regions:
%
\begin{align*}
& (\vect{x}^L,\vect{x}^U) \bullet (\vect{y}^L,\vect{y}^U) \\
= &
\begin{cases}
(\vect{x}^L, \vect{y}^U) & \vect{x}^U_1 + 1 = \vect{y}^L_1 \; \wedge \;
(\vect{x}^L_{1:n}, \vect{x}^U_{1:n}) = (\vect{y}^L_{1:n}, \vect{y}^U_{1:n}) \\
\bot  & \textit{otherwise}
\end{cases}
\end{align*}
For our example, this yields:
%
\begin{align*}
Q(v) = \{&(\pi^2_1, [
\stwo{0}{-1}{0}{-1}
\stwo{-1}{0}{1}{0}, 
\stwo{0}{1}{0}{1}, 
]) \\
&(\pi^2_2, [
\stwo{0}{-1}{0}{-1},
\stwo{-1}{0}{1}{0},
\stwo{0}{1}{0}{1}])\}
\end{align*}
%
\item Unpermute and union together, \ie{},
%
\[
U(v) = \bigcup \{[\pi \vect{y}^L, \pi \vect{y}^U]
 \mid [\vect{y}^L, \vect{y}^U] \leftarrow S, (\pi, S) \leftarrow Q(v)\}
\]
Which yields:
%
\begin{align*}
U(\texttt{a}) = 
\{\stwo{0}{-1}{0}{-1}
\stwo{-1}{0}{1}{0}
\stwo{0}{1}{0}{1}
\stwo{-1}{0}{-1}{0}
\stwo{0}{-1}{0}{1}
\stwo{1}{0}{1}{0}\}
\end{align*}
%
\item Finally, filter by the region containment predicate $\containedin$, that
  is, if any region is contained within another then remove the
  smaller. The $\containedin$ predicate is defined:
%
\begin{align*}
& (\vect{x}^L, \vect{x}^U) \containedin (\vect{y}^L, \vect{y}^U) = \\
& \vect{y}^L_1 \leq \vect{x}^L_1 \wedge \vect{x}^U_1 \leq \vect{y}^U_1
  \wedge (\vect{x}^L_{1:n}, \vect{x}^U_{1:n}) \containedin
  (\vect{y}^L_{1:n}, \vect{y}^U_{1:n})
\end{align*}
For our example this then yields the final result of: 
\begin{align*}
\textsf{regions}(M) 
= \{\stwo{-1}{0}{1}{0} \stwo{0}{-1}{0}{1}\}
\end{align*}
since $\stwo{0}{1}{0}{1} \sqsubseteq \stwo{0}{-1}{0}{1}$
and $\stwo{1}{0}{1}{0} \sqsubseteq \stwo{-1}{0}{1}{0}$. 
\end{enumerate}
For our example, applying \textsf{regions} again yields the same
result, hence we have reached a fixed point.

>>>>

\newcommand{\stenFwd}[2]{\texttt{forward} \, (\texttt{depth=}#1,
  \texttt{dim=}#2)}
\newcommand{\stenBwd}[2]{\texttt{backward} \, (\texttt{depth=}#1,
  \texttt{dim=}#2)}
\newcommand{\stenCen}[2]{\texttt{centered} \, (\texttt{depth=}#1,
  \texttt{dim=}#2)}
\newcommand{\stenFwdS}[2]{\texttt{fwd} \, (\texttt{depth=}#1,
  \texttt{dim=}#2)}
\newcommand{\stenBwdS}[2]{\texttt{bwd} \, (\texttt{depth=}#1,
  \texttt{dim=}#2)}
\newcommand{\stenCenS}[2]{\texttt{cen} \, (\texttt{depth=}#1,
  \texttt{dim=}#2)}

%% VECTOR NOTATIONS
\newcommand{\vect}[1]{\textbf{#1}}
\newcommand{\vtwo}[2]{\setlength{\arraycolsep}{0em}
\left[\begin{array}{l}$#1$\\$#2$\end{array}\right]}
\newcommand{\vthree}[3]{\setlength{\arraycolsep}{0em}
\left[\begin{array}{l}$#1$\\$#2$\\$#3$\end{array}\right]}
\newcommand{\stwo}[4]
%{\vtwo{#1}{#2}\!\vtwo{#3}{#4}}
{\setlength{\arraycolsep}{0.1em}
\left[\begin{array}{rr}$#1$ & $#3$\\$#2$ & $#4$\end{array}\right]}

\newcommand{\singleEntry}[2]{\textbf{J}_{#2}^{#1}}

@@

\begin{figure}[t]
\begin{align*}
\def\arraystretch{1.2}
\setlength{\arraycolsep}{0.2em}
\newcommand{\dimTy}{\mathbb{D}}
\begin{array}{rl}
\nonterm{specification} ::= & \nonterm{regionDec} \mid \nonterm{specDec} \\
\nonterm{specDec} ::= & \term{stencil} \; \nonterm{spec}
  \\
\nonterm{regionDec} ::= &  \texttt{region} \; \nonterm{rvar} \; \texttt{=} \;
                         \nonterm{region}\\
\nonterm{spec} ::= & \nonterm{spatial} \mid \nonterm{temporal}   \\[1em]
\nonterm{spatial} ::= & [\nonterm{mod}] \;
                       [\nonterm{approxMod},] \; [\nonterm{region}] \\
\nonterm{mod} ::= & [\nonterm{modR},] \; [\nonterm{modRefl},] \; [\nonterm{modIrefl},] \\
\nonterm{modR} ::= &
 \term{readOnce} \\
\nonterm{modRefl} ::= & \term{reflexive}, \; \term{dims=}\dimTy \{\term{,} \; \dimTy\} \\
\nonterm{modIrrefl} ::= & \term{irreflexive}, \; \term{dims=}\dimTy \{\term{,}
          \dimTy\}\\
\nonterm{approxMod} ::= & \term{atMost} \; \mid \; \term{atLeast} \\[0.1em]
\nonterm{region} ::= & \term{forward}, \term{depth=}\mathbb{N}_{>0}, \term{dim=}\dimTy  \\
\mid \; & \term{backward}, \term{depth=}\mathbb{N}_{>0}, \term{dim=}\dimTy \\
\mid \; & \term{centered}, \term{depth=}\mathbb{N}_{>0}, \term{dim=}\dimTy \\
\mid \; & \nonterm{region} \; \term{+} \; \nonterm{region} \\
\mid \; & \nonterm{region} \; \term{*} \; \nonterm{region} \\
\mid \; & \nonterm{rvar}  \\[0.5em]
\nonterm{temporal} ::= \; & \term{dependency} \; (v \; \{ , v \}) [, \texttt{mutual}]
  \\[0.5em]
\dimTy ::= \; & \mathbb{N}_{>0} \\
\nonterm{rvar} ::= \; & [\text{\term{a}-\term{z}$\,$\term{A}-\term{Z}$\,$\term{0}-\term{9}}]+
\end{array}
\end{align*}
\caption{Specification syntax (EBNF grammar)}
\label{fig:syntax}
\end{figure}

@@

\begin{figure}
\begin{align*}
\setlength{\arraycolsep}{0.05em}
\begin{array}{c}
\framebox{$\textit{region} \equiv \textit{region}'$} \\[1em]
\begin{array}{rll}
(\textsc{F\; \texttt{+} \;F}) \;\; &
\stenFwdS{n \, \textsf{max} \, m}{d} \\
 \equiv \; & \stenFwdS{n}{d} \; \texttt{+} \; \stenFwdS{m}{d} \\[1em]
%
(\textsc{B\; \texttt{+} \;B}) \;\; &
\stenBwdS{n \, \textsf{max} \, m}{d} \\
 \equiv \; & \stenBwdS{n}{d} \; \texttt{+} \; \stenBwdS{m}{d} \\[1em]
%
(\textsc{C\; \texttt{+} \;C}) \;\; &
\stenCenS{n \, \textsf{max} \, m}{d} \\
\equiv \; & \stenCenS{n}{d} \; \texttt{+} \; \stenCenS{m}{d} \\[1em]
%
(\textsc{C\; \texttt{+} \;F}) \;\; & \stenCenS{n}{d} \\
\textit{$m \leq n$} \; \equiv \; & \stenCenS{n}{d} \; \texttt{+} \;
                      \stenFwdS{m}{d} \\[1em]
%
(\textsc{C\; \texttt{+} \;B}) \;\; & 
\stenCenS{n}{d} \\
\textit{$m \leq n$} \; \equiv \; & \stenCenS{n}{d} \; \texttt{+} \;
                      \stenBwdS{m}{d} \\[1em]
(\textsc{B\; \texttt{+} \;F}) \;\; & 
\stenCenS{n}{d} \\
\equiv \; & \stenFwdS{n}{d} \; \texttt{+} \; \stenBwdS{n}{d} \\[1em]
%
(\texttt{+}\textsc{comm}) \;\; & S \; \texttt{+} \; T \; \equiv \; T \;
                       \texttt{+} \; S \\[0.5em]
%
(\texttt{+}\textsc{assoc}) \;\; & R \; \texttt{+} \; (S \; \texttt{+} \; T) \; \equiv \; (R \;
                       \texttt{+} \; S) \; \texttt{+} \; T \\[0.5em]
(\texttt{*}\textsc{comm}) \;\; & S \; \texttt{*} \; T \; \equiv \; T \;
                       \texttt{*} \; S \\[0.5em]
%
(\texttt{*}\textsc{assoc}) \;\; & R \; \texttt{*} \; (S \; \texttt{*} \; T) \; \equiv \; (R \;
                       \texttt{*} \; S) \; \texttt{*} \; T \\[0.5em] 
(\textsc{dist}) \;\; & R \; \texttt{*} \; (S \; \texttt{+} \; T) \; \equiv \; (R \;
                       \texttt{*} \; S) \; \texttt{+} \; (R
                       \; \texttt{*} \; T) \\[0.5em] 
(\textsc{inter}) \;\; & (S \; \texttt{+} \; S') \; \texttt{*} \;
                              (T \; \texttt{+} \; T') \\
 \equiv \; & (S \;  \texttt{*} \; S) 
 \; \texttt{+} \; (S \; \texttt{*} \; T') 
 \; \texttt{+} \; (S' \; \texttt{*} \; T)
 \; \texttt{+} \; (S' \; \texttt{*} T')  \\[0.5em] 
\end{array} \\ \\
%
\framebox{$\overline{\textit{specDec} :: \texttt{v}} \equiv 
\overline{\textit{specDec'} :: \texttt{v}}$} \\[1em]
\begin{array}{rl}
(\textsc{mutual}) \; & 
\texttt{stencil} \; \texttt{dependency} (\texttt{v}), \texttt{mutual}
  :: \texttt{w} 
\\
\equiv \; & \texttt{stencil} \; \texttt{dependency} (\texttt{w}), \texttt{mutual} ::
  \texttt{v} \\[0.5em]
(\textsc{coalesce}) \; & 
\texttt{stencil} \; \texttt{dependency} (\bar{v}) 
  :: \texttt{v} \\
& ; \texttt{stencil} \; \texttt{dependency} (\bar{w}) 
  :: \texttt{v} 
\\
\equiv \; & \texttt{stencil} \; \texttt{dependency} (\bar{v}, \bar{w}) ::
  \texttt{v} \\[0.5em]
(\textsc{spatial}) \; &
(\textit{region} \equiv \textit{region'}) \\
\Rightarrow \; & (\texttt{stencil} \; [\textit{mod}] \; [\textit{approxMod},] \;
\textit{region} \\
& \!\!\equiv \; \texttt{stencil} \; [\textit{mod}] \;
            [\textit{approxMod},] \;
\textit{region}') \\
\end{array}
\end{array}
\end{align*}
\caption{Equations on specifications}
\label{fig:equations}
\end{figure}

@@

@@

\item Sort each permutation set into an ordered list, based on the ordering:
\[
\vect{x} \leq \vect{y} = \exists i \, . \; \vect{x}_i \leq \vect{y}_i 
\; \wedge \; \vect{x}_{i:n} = \vect{y}_{i:n}
\]
%
This yields:
%%
\begin{align*}
\{&(\pi^2_1, [
\stwo{0}{-1}{0}{-1}
\stwo{-1}{0}{-1}{0} 
\stwo{0}{0}{0}{0}
\stwo{1}{0}{1}{0} 
\stwo{0}{1}{0}{1}] )
\\
&(\pi^2_2, [
\stwo{0}{-1}{0}{-1}
\stwo{-1}{0}{-1}{0}
\stwo{0}{0}{0}{0}
\stwo{1}{0}{1}{0}
\stwo{0}{1}{0}{1}])\}
\end{align*}
%%
\item Fold each list pairwise by the following partial operation
 $\bullet$ which coalescees contiguous regions:
%
\begin{align*}
& (\vect{x}^L,\vect{x}^U) \bullet (\vect{y}^L,\vect{y}^U) \\
= &
\begin{cases}
(\vect{x}^L, \vect{y}^U) & \vect{x}^U_1 + 1 = \vect{y}^L_1 \; \wedge \;
(\vect{x}^L_{1:n}, \vect{x}^U_{1:n}) = (\vect{y}^L_{1:n}, \vect{y}^U_{1:n}) \\
\bot  & \textit{otherwise}
\end{cases}
\end{align*}
For our example, this yields:
%
\begin{align*}
Q(v) = \{&(\pi^2_1, [
\stwo{0}{-1}{0}{-1}
\stwo{-1}{0}{1}{0}, 
\stwo{0}{1}{0}{1}, 
]) \\
&(\pi^2_2, [
\stwo{0}{-1}{0}{-1},
\stwo{-1}{0}{1}{0},
\stwo{0}{1}{0}{1}])\}
\end{align*}
%
\item Unpermute and union together, \ie{},
%
\[
U(v) = \bigcup \{[\pi \vect{y}^L, \pi \vect{y}^U]
 \mid [\vect{y}^L, \vect{y}^U] \leftarrow S, (\pi, S) \leftarrow Q(v)\}
\]
Which yields:
%
\begin{align*}
U(\texttt{a}) = 
\{\stwo{0}{-1}{0}{-1}
\stwo{-1}{0}{1}{0}
\stwo{0}{1}{0}{1}
\stwo{-1}{0}{-1}{0}
\stwo{0}{-1}{0}{1}
\stwo{1}{0}{1}{0}\}
\end{align*}
%
\item Finally, filter by the region containment predicate $\containedin$, that
  is, if any region is contained within another then remove the
  smaller. The $\containedin$ predicate is defined:
%
\begin{align*}
& (\vect{x}^L, \vect{x}^U) \containedin (\vect{y}^L, \vect{y}^U) = \\
& \vect{y}^L_1 \leq \vect{x}^L_1 \wedge \vect{x}^U_1 \leq \vect{y}^U_1
  \wedge (\vect{x}^L_{1:n}, \vect{x}^U_{1:n}) \containedin
  (\vect{y}^L_{1:n}, \vect{y}^U_{1:n})
\end{align*}
For our example this then yields the final result of: 
\begin{align*}
\textsf{regions}(M) 
= \{\stwo{-1}{0}{1}{0} \stwo{0}{-1}{0}{1}\}
\end{align*}
since $\stwo{0}{1}{0}{1} \sqsubseteq \stwo{0}{-1}{0}{1}$
and $\stwo{1}{0}{1}{0} \sqsubseteq \stwo{-1}{0}{1}{0}$. 
\end{enumerate}
For our example, applying \textsf{regions} again yields the same
result, hence we have reached a fixed point.
